Correlation between different emission channels from an atom inside a leaky driven cavity

### 2019-12-17 newpaper.tex
algunos cambios de pablo

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.2 distribution.
%   Version 4.2a of REVTeX, December 2014
%
%   Copyright (c) 2014 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.2
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
\documentclass[%
 reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps, 
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-2}

@@

\usepackage{physics}
\usepackage{fixme}
%\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

@@

\section{INTRODUCTION}

Spontaneous emission is a phenomenon that has been studied since the
early XX century \cite{10.2307/94746, 1917PhyZ...18..121E}. One of the
main characteristics of spontaneous emission is that
\fxnote{spontaneus emission no es exponencial, es aproximadamente
  exponencial, si quitas la aproximacion markoviana te queda una cola
  polinomial, quite la oracion} the exact time of emission, as well as
the direction in which the photon is emitted, is random, which
difficults its measurement. On the other hand, photons emitted from a
cavity are easy to measure, partly due the fact that we know that they
leave the cavity through one of the lateral mirrors \cite{326305,
  doi:10.1063/1.113345}. The randomness of the direction in which the
photons are emitted by atoms in free space difficults photon counting
experiments. To sort out this difficulty one could think of a physical
arrangement, in which the two-level atom is coupled to the stimulated
mode of a cavity, as it is shown in Fig. \ref{asa} (b). In this
physical arrangement, \fxnote*{tienes una cita donde lo nombren
  asi?}{which is known as the driven Jaynes-Cummings model}, one could
try to estimate the number of spontaneous emissions of the atom
--which are emitted in modes that are not cavity modes and therefore
escape the cavity-- by measuring the photons that leak through one of
the cavity mirrors. In the case that a
correlation exists between the two quantities, we can count the number
of photons that leak the cavity through its mirror, which is easy to
measure, to find information about the number of photons emitted into
the other modes (spontaneus emission), which are difficult to measure.

In the case of a two-level atom in free space, the absence of
correlation between different channels is the result of the markovian
approximation, which assumes that the environment has no memory of
previous interactions with the system \cite{daley2014quantum}. Two
atomic emissions would be correlated if one of the emissions stay
registered in the environment and, somehow, influence the second
emission of a two-level atom at a later time; the Markovian
approximation does not allow memory effects and there is no
correlation between the number of emissions of each source: knowing
the number of photons emitted in some region of the space during some
time does not give me any information on the number of photons emitted
in other region of the space. Nevertheless, for the case of an atom
inside a cavity, before a photon leaks the cavity, it will interact
with the atom for some time. This interaction with the atom before the
emission generates a correlation between photons that leak the cavity
through its mirror and photons that leak the cavity through its side.

We use quantum trajectory theory to theoretically calculate the
correlation between different emission channels. A quantum trajectory
is the description of a continuously monitored quantum system
conditioned on a particular past history of coherent evolution and
collapse \cite{Carmichael1993Open}. These trajectories are simulated
through a stochastic process, which define the times when the
collapses occur. Each trajectory gives us the number of photon emitted
in each channel. Simulating a large number of trajectories and
registering the number of photon emissions into the two channels, we
obtain its correlation. The correlation varies with the cavity
linewidth and the atom-cavity coupling. For a very bad cavity we
recover the free space result: there is no correlation between the two
channels.

This article is organized as follows: A general overview of the driven
Jaynes-Cummings model is presented in section \ref{sc:drivenjc}. In
section \ref{sc:distributions} we obtain probability distributions
that relate the number of spontaneous and cavity emissions. In section
\ref{sc:correlation} we calculate the correlation between the two variables.
Concluding remarks are given in section \ref{sc:conclusions}
\begin{center}
\begin{figure*}\label{asa}
\begin{center}
\includegraphics[scale = 0.65]{newimagepaper.pdf}
%\includegraphics[scale = 0.45]{image_paper222.png}
\caption{\small{Physical arrangements of atom and atom in a cavity. In
    both of these arrangements we consider a source which stimulates
    the atom provoking it to spontaneously emit, once it decays. We
    consider a detector in two directions, and we keep a register of
    the number of photons measured in each direction. (a) In the case
    for the atom we don't expect a correlation between m (the number
    of photons measured in the upper side of the atom) and n (the
    number of photons measured in the right side of the atom); (b) for
    the atom in the cavity (a cavity with some loss rate) we study the
    possibility that the cavity generates a correlation between the
    emissions of the atom and the emissions of the cavity, which
    correspond to the emissions that leak through mirror
    2.}} 
\end{center}
\end{figure*}
\end{center}

\section{Driven Jaynes-Cummings model}\label{sc:drivenjc}
e consider a system consisting of a two-level atom, coupled to a mode
of a leaky cavity. The mode is driven by a coherent field resonant
with the atomic transition and the cavity mode.
% In order to account for
% dissipation one may use a non-hermitian Hamiltonian to describe the
% system \cite{bla}. Using this description for the system will turn
% useful at the moment of using quantum trajectory theory. Such
% Hamiltonian can be separated into hermitian and non-hermitian terms.
%\begin{equation}
%H = H_s - \frac{i\hbar}{2}\sum_m C^\dagger_n C_n .
%\end{equation}
The Hamiltonian of this system, in the rotating wave approximation is
given by
\begin{equation}
H_s = H_{at} + H_{cav} + H_{atcav} + H_{lascav}, \label{mainham}  
\end{equation}
where the atom and cavity terms are, respectively
\begin{subequations}
\begin{equation}
H_{at} = \tfrac{1}{2}\hbar \omega \sigma_z,    
\end{equation}
\begin{equation}
H_{cav} = \hbar \omega  a^\dagger a,  
\end{equation}
while the coupling between atom and cavity, and cavity and coherent
field are
\begin{equation}
H_{atcav} = i\hbar g(a\sigma_+ - a^\dagger \sigma_-),    
\end{equation}
\begin{equation}
H_{lascav} = i\hbar \mathcal{E}(ae^{i\omega t} - a^\dagger e^{-i\omega t}),    
\end{equation}
\end{subequations}
where the annihilation operator $a$ of a single photon in the cavity,
and the creation operator $a^\dagger$ obey $[a, a^\dagger] = 1$. The
operators $\sigma_-, \sigma_+$ account for the atomic lowering and
raising operators, respectively, and the operator $\sigma_z$ accounts
for the population difference between the atomic levels. These
operators obey
$[\sigma_+, \sigma_-] = \sigma_z, \ [\sigma_\pm, \sigma_z] = \mp
2\sigma_\pm$. The strength of the coupling between atom and cavity
field is given by the constant
\begin{equation}
g = \sqrt{\frac{\omega d^2}{2\hbar \epsilon_0 V_Q}},
\end{equation}
where $V_Q$ is the mode volume , $d$ is the atomic dipole , and
$\epsilon_0$ is the permitivity of free space. $\mathcal{E}$ is a term
proportional to the coherent field amplitude. Dissipation is
introduced by the following non-hermitian Hamiltonian
\begin{equation} \label{nh}
H_{nh} = H_{dec} + H_{leak},
\end{equation}
where the terms corresponding to spontaneous emission of the atom and cavity losses are
\begin{subequations}
\begin{equation}
H_{dec} = - i\hbar\frac{\gamma}{2}\sigma_+\sigma_-,
\end{equation}
\begin{equation}
H_{leak} = - i\hbar\kappa a^\dagger a\, ,
\end{equation}
\end{subequations}
where $\kappa$ and $\gamma$ are the cavity loss and spontaneous decay
rate. The evolution between jumps is given by $H=H_s+H_{nh}$. This
evolution is interrumped by spontaneus emission by the atom or when a
photon leaks the cavity.

>>>>

% ****** Start of file apssamp.tex ******
%
%   This file is part of the APS files in the REVTeX 4.2 distribution.
%   Version 4.2a of REVTeX, December 2014
%
%   Copyright (c) 2014 The American Physical Society.
%
%   See the REVTeX 4 README file for restrictions and more information.
%
% TeX'ing this file requires that you have AMS-LaTeX 2.0 installed
% as well as the rest of the prerequisites for REVTeX 4.2
%
% See the REVTeX 4 README file
% It also requires running BibTeX. The commands are as follows:
%
%  1)  latex apssamp.tex
%  2)  bibtex apssamp
%  3)  latex apssamp.tex
%  4)  latex apssamp.tex
%
\documentclass[%
 reprint,
%superscriptaddress,
%groupedaddress,
%unsortedaddress,
%runinaddress,
%frontmatterverbose, 
%preprint,
%preprintnumbers,
%nofootinbib,
%nobibnotes,
%bibnotes,
 amsmath,amssymb,
 aps, 
%pra,
%prb,
%rmp,
%prstab,
%prstper,
%floatfix,
]{revtex4-1}

@@

\usepackage{physics}
\usepackage[draft,inline,nomargin]{fixme} \fxsetup{theme=color}
%\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
%\usepackage{hyperref}% add hypertext capabilities
%\usepackage[mathlines]{lineno}% Enable numbering of text and display math
%\linenumbers\relax % Commence numbering lines

@@

\section{Introduction}

Spontaneous emission is a phenomenon that has been studied since the
early XX century \cite{10.2307/94746, 1917PhyZ...18..121E}. One of the
main characteristics of spontaneous emission is that the exact time of
emission, as well as the direction in which the photon is emitted, is
random, which difficults its measurement. On the other hand, photons
emitted from a cavity mirror are easier to measure because we know
their direction \cite{326305, doi:10.1063/1.113345}. The randomness of
the direction in which the photons are emitted by atoms in free space
difficults photon counting experiments. To sort out this difficulty
one could think of a physical arrangement in which a two-level atom is
coupled to a driven cavity, as it is shown in Fig. \ref{asa}. In this
physical arrangement, one could try to estimate the number of
spontaneous emissions of the atom --which are emitted in modes that
are not cavity modes and therefore escape the cavity to its side-- by
measuring the photons that leak through one of the cavity mirrors. In
the case that a correlation exists between the two quantities, we can
count the number of photons that leak the cavity through its mirror,
which is easy to measure, to find information about the number of
photons emitted into the other modes (spontaneus emission), which are
difficult to measure.

In the case of a two-level atom in free space, the absence of
correlation between photons emitted in different directions is due to
the fact that the atom's environment memory of previous interactions
with it, is very short; this is what justify the markovian
approximation \cite{daley2014quantum}. Two atomic emissions would be
correlated if one of the emissions stay registered in the environment
and, somehow, influence the second emission.

For an atom in free space, knowing the number of photons emitted in
some region of the space during some time does not give any
information on the number of photons emitted in other region of the
space. Nevertheless, for the case of an atom inside a cavity, before a
photon, emitted from the atom into the cavity mode, leaks the cavity,
it will interact with the atom for some time. This interaction creates
a correlation between photons that leak the cavity through its mirror
and photons that are emitted from the atom to the cavity side
(spontaneus emissions).

We use quantum trajectories theory to calculate the correlation between
different emission channels. A quantum trajectory is the description
of a continuously monitored quantum system conditioned on a particular
past history of coherent evolution and collapse
\cite{Carmichael1993Open}. These trajectories are simulated through a
stochastic process, which define the times when the collapses occur.
Each trajectory gives us the number of photon emitted in each channel.
Simulating a large number of trajectories and registering the number
of photon emissions into the two channels, we obtain its correlation.
The correlation varies with the cavity linewidth and the atom-cavity
coupling. For a very bad cavity we recover the free space result:
there is no correlation between the two channels.

This article is organized as follows: A general overview of the
theoretical model is presented in section \ref{sc:drivenjc}. In
section \ref{sc:distributions} we obtain join probability distribution
of the the number of spontaneous and cavity emissions. In section
\ref{sc:correlation} we calculate the correlation between the two
variables. Concluding remarks are given in section
\ref{sc:conclusions}
\begin{center}
\begin{figure*}\label{asa}
\begin{center}
\includegraphics[scale = 0.65]{newimagepaper.pdf}
%\includegraphics[scale = 0.45]{image_paper222.png}
\caption{A driven atom in free space (a) and inside a driven leaky
  cavity (b). Each time the atom decays, it emits a photon. In each
  case, we count the number of photons $m$ and $n$, emitted in two
  different directions respectively. For the atom in free space there
  is no correlation between $m$ and $n$. For the atom inside the leaky
  cavity we find a correlation between $m$ and $n$.}
\end{center}
\end{figure*}
\end{center}

\section{Theoretical model}\label{sc:drivenjc}
We consider a system consisting of a two-level atom, coupled to a mode
of a leaky cavity. The mode is driven by a coherent field resonant
with the atomic transition and the cavity mode.
% In order to account for
% dissipation one may use a non-hermitian Hamiltonian to describe the
% system \cite{bla}. Using this description for the system will turn
% useful at the moment of using quantum trajectory theory. Such
% Hamiltonian can be separated into hermitian and non-hermitian terms.
%\begin{equation}
%H = H_s - \frac{i\hbar}{2}\sum_m C^\dagger_n C_n .
%\end{equation}
The Hamiltonian of this system, in the rotating wave approximation is
given by
\begin{equation}
H_s = H_{at} + H_{cav} + H_{atcav} + H_{lascav}, \label{mainham}  
\end{equation}
where the atom and cavity terms are, respectively
\begin{subequations}
\begin{equation}
H_{at} = \tfrac{1}{2}\hbar \omega \sigma_z,    
\end{equation}
\begin{equation}
H_{cav} = \hbar \omega  a^\dagger a,  
\end{equation}
while the coupling between atom and cavity, and cavity and coherent
field are
\begin{equation}
H_{atcav} = i\hbar g(a\sigma_+ - a^\dagger \sigma_-),    
\end{equation}
\begin{equation}
H_{lascav} = i\hbar \mathcal{E}(ae^{i\omega t} - a^\dagger e^{-i\omega t}),    
\end{equation}
\end{subequations}
where the annihilation operator $a$ of a single photon in the cavity,
and the creation operator $a^\dagger$ obey $[a, a^\dagger] = 1$. The
operators $\sigma_-, \sigma_+$ account for the atomic lowering and
raising operators, respectively, and the operator $\sigma_z$ accounts
for the population difference between the atomic levels. These
operators obey
$[\sigma_+, \sigma_-] = \sigma_z, \ [\sigma_\pm, \sigma_z] = \mp
2\sigma_\pm$. The strength of the coupling between atom and cavity
field is given by the constant
\begin{equation}
g = \sqrt{\frac{\omega d^2}{2\hbar \epsilon_0 V_Q}},
\end{equation}
where $V_Q$ is the mode volume , $d$ is the atomic dipole , and
$\epsilon_0$ is the permitivity of free space. $\mathcal{E}$ is a term
proportional to the coherent field amplitude. Dissipation is
introduced by the following non-hermitian Hamiltonian
\begin{equation} \label{nh}
H_{nh} = H_{dec} + H_{leak},
\end{equation}
where the terms corresponding to spontaneous emission of the atom and cavity losses are
\begin{subequations}
\begin{equation}
H_{dec} = - i\hbar\frac{\gamma}{2}\sigma_+\sigma_-,
\end{equation}
\begin{equation}
H_{leak} = - i\hbar\kappa a^\dagger a\, ,
\end{equation}
\end{subequations}
where $\kappa$ and $\gamma$ are the cavity loss and spontaneous decay
rate. The evolution between jumps is given by $H=H_s+H_{nh}$. This
evolution is interrumped by spontaneus emission by the atom or when a
photon leaks the cavity.
